---
title: "Biaxial sleepwalk test"
output: html_document
date: "2024-04-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r functions}
setwd(here::here("src"))
source("knn_sleepwalk.R")
```

Here, we're going to test a modification of sleepwalk that compares a biaxial plot to its corresponding region on a UMAP. Let's load the CyTOF data.

```{r get_data}
library(tidyverse)
library(here)
library(flowCore)
library(sleepwalk)
set.seed(1)

setwd(here::here('data'))

# Read in the cells
cells <- flowCore::read.FCS(list.files(pattern = "SLE")) # Data from Marie Burns
params <- as.vector(Biobase::pData(parameters(cells))$desc)
colnames(cells) <- params
cells <- exprs(cells) 
cells <- cells[,!is.na(colnames(cells))]
cells <- as_tibble(cells)

# Filter the cells by marker we're going to use
marker_info <- readr::read_csv("cytof_marker_data.csv")
marker_info <- dplyr::filter(marker_info, is.na(notes)) # Take this out later

# Subsample
num_cells <- 10000
cells <- cells[sample(nrow(cells), num_cells),]
cells

# Tranform the markers we want to transform
to_transform <- dplyr::filter(marker_info, marker_type != "none")$desc
keep_same <- dplyr::filter(marker_info, marker_type == "none")$desc %>% .[!is.na(.)]

tmp1 <- cells[,to_transform]
tmp2 <- cells[,keep_same]

tmp1 <- asinh(tmp1/5)
cells <- bind_cols(tmp1, tmp2)
```

Ok, now let's make our UMAP.

```{r run_dimr}
library(umap)

surface <- cells[,dplyr::filter(marker_info, marker_type == "type")$desc]
umap <- umap::umap(surface)$layout %>% as_tibble()
umap
```
We're now going to modify the marker names to make it easier to do this next part.

```{r}
NumberDuplicates <- function(x) {
  # Count occurrences and make adjustments if necessary
  ux <- unique(x)
  for (i in ux) {
    # Indices of each unique string
    indices <- which(x == i)
    if (length(indices) > 1) {
      # Modify elements at those indices
      x[indices] <- paste0(i, seq_along(indices))
    }
  }
  return(x)
}

# Take care of outlier
if(names(cells)[19] == "142Nd_cleaved_caspase_3") {
   names(cells)[19] <- "142Nd_cleaved-caspase-3"
} else {
   stop("please check the caspace 3 column")
}

# Make the naming easier
names(cells) <- sub(".*_", "", names(cells)) %>% NumberDuplicates()
names(cells)
```


And now for the sleepwalk hack.

```{r}
setwd(here::here("output"))

biaxial <- bind_cols(cells$CD3, cells$CD19)

KnnSleepwalk(mat = biaxial, embedding = list(biaxial, umap), k = 50, output_file = paste0("knn_sleepwalk_biaxial_test", "_", num_cells, ".html"), point_size = 4, dimr_names = c("", ""))
```
Now we are going to do multiple biaxials along with the UMAP. We won't worry about the killer app for back gating just yet. Let's just see what it looks like.

```{r}
setwd(here::here("output"))

cd3_cd19 <- bind_cols(cells$CD3, cells$CD19)
cd4_cd8 <- bind_cols(cells$CD4, cells$CD8)
cd4_cd45ro <- bind_cols(cells$CD4, cells$CD45RO)
cd45ro_cd45ra <- bind_cols(cells$CD45RO, cells$CD45RA)
dna1_dna2 <- bind_cols(cells$DNA1, cells$DNA2)
cd45_cd66b <- bind_cols(cells$CD45, cells$CD66b)
cd3_cd11b <- bind_cols(cells$CD3, cells$CD11b)

# KnnSleepwalk(mat = cd45ro_pnfkb, embedding = list(cd45ro_pnfkb, cd4_cd45ro, cd4_cd8, cd3_cd19, umap), k = 50, output_file = paste0("knn_sleepwalk_biaxial_backgating", "_", num_cells, ".html"), point_size = 4, dimr_names = c("", ""))

KnnSleepwalk(mat = cd45ro_cd45ra, embedding = list(cd45ro_cd45ra, cd4_cd8, cd3_cd19, cd45_cd66b, umap), k = 50, output_file = paste0("knn_sleepwalk_biaxial_backgating", "_", num_cells, ".html"), point_size = 4, dimr_names = c("", "", "", "", ""))
```
At this point, we have to actually simulate back gating. How do we do that? By setting non-gated cells to zero. Let's redo the above example.

```{r}
setwd(here::here("output"))

# Some gating, using the origin point solution for hierarchies
cd45_gate <- apply(cells, 2, function(i) {
    ifelse(cells$CD45 > 4 & cells$CD66b < 2, i, i + 10)
}) %>% as_tibble()

cd3_gate <- apply(cd45_gate, 2, function(i) {
    ifelse(cd45_gate$CD3 > 3, i, i + 10)
}) %>% as_tibble()

cd4_gate <- apply(cd3_gate, 2, function(i) {
    ifelse(cd3_gate$CD4 > 2 & cd3_gate$CD8 < 3, i, i + 10)
}) %>% as_tibble()

cd3_cd19 <- bind_cols(cd45_gate$CD3, cd45_gate$CD19)
cd4_cd8 <- bind_cols(cd3_gate$CD4, cd3_gate$CD8)
cd45ro_cd45ra <- bind_cols(cd4_gate$CD45RO, cd4_gate$CD45RA)
cd45_cd66b <- bind_cols(cells$CD45, cells$CD66b)

KnnSleepwalk(mat = cd45ro_cd45ra, embedding = list(cd45ro_cd45ra, cd4_cd8, cd3_cd19, cd45_cd66b, umap), k = 50, output_file = paste0("knn_sleepwalk_biaxial_backgating2", "_", num_cells, ".html"), point_size = 4, dimr_names = c("", "", "", "", ""))
```

