---
title: "Inverted umap test"
output: html_document
date: "2024-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Inspired by information I got from the sleepwalk experiments in this project, I am going to try to build a UMAP with variations of the distance matrix. Particularly the inverted distance matrix. First, we load up everything.

```{r process_data}
library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)

# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "data/filtered_gene_bc_matrices/hg19/")
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)

# Processing
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(pbmc), 10)
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
```

And now we're going to run the UMAP.

```{r}
library(umap)

pca <- pbmc@reductions$pca@cell.embeddings %>% as_tibble()
umap_reg <- umap::umap(pca)$layout %>% as_tibble()
names(umap_reg) <- c("umap1", "umap2")
```

```{r}
ggplot(umap_reg, aes(x = umap1, y = umap2)) + geom_point()
```

Great. And now for the big experiment. First we run it on the distance matrix, which should in theory work.

```{r}
pca_dist <- dist(pca) %>% as.matrix()
umap_dist <- umap::umap(pca_dist, input = "dist")$layout %>% as_tibble()
names(umap_dist) <- c("umap1", "umap2")
```

```{r}
ggplot(umap_dist, aes(x = umap1, y = umap2)) + geom_point()
```

Ok, that worked. Now we invert.

```{r}
# Add 0.01 to make sure nothing is divided by 0
umap_dist_inverted <- umap::umap(1/(pca_dist + 0.01), input = "dist")$layout %>% as_tibble()
names(umap_dist_inverted) <- c("umap1", "umap2")
```

```{r}
ggplot(umap_dist_inverted, aes(x = umap1, y = umap2)) + geom_point()
```

Ok, this is weird. Now for our sleepwalk.

```{r functions}
source('src/knn_sleepwalk.R')
```


```{r}
setwd(here::here("output"))

KnnSleepwalkDirect(mat1 = umap_dist, mat2 = pca, embedding = list(umap_dist, umap_dist), k = 27, point_size = 4, output_file = "umap_dist_knn_sleepwalk_scrna_seq_umap_vs_pca_space.html")

KnnSleepwalkDirect(mat1 = umap_dist_inverted, mat2 = pca, embedding = list(umap_dist_inverted, umap_dist_inverted), k = 27, point_size = 4, output_file = "umap_dist_inverted_knn_sleepwalk_scrna_seq_umap_vs_pca_space.html")
```

